---
output:
  word_document: default
  pdf_document: default
  html_document: default
---
# Why fake death? Environmental and genetic control of tonic immobility in larval lacewings 

```{r, include=FALSE}
require(knitr)
require(plyr)
require(dplyr)
require(tidyr)
library(rcompanion)
require(ggplot2)
require(car)
require(FSA)
require(scales)
require(MCMCglmm)
require(knitr)
library(emmeans)
theme_set(theme_classic(base_size=18))
```


```{r}
dat <- read.csv(file="~/Documents/Death_Feign_Paper/Tonic_Immobility_5.18.23/DeathFeignAllData.csv", header=TRUE)
```

## Analysis of head and body measurements

Subset data to exclude offspring from crosses

```{r}
pardat <- subset(dat, dat$generation != "O")
pardat$diet <- as.factor(pardat$diet)
```

## Plots of head and body measures

The effect of treatment on body size. Multiple peaks in head capsule size indicate multiple instars.

```{r, warning=FALSE}
ggplot(pardat, aes(x = head_length * 25.4, group = factor(diet), colour = factor( diet))) + 
  geom_density() +  
  guides(alpha = FALSE) +
  scale_colour_manual(name = "",
                      labels = c("Fed","1 Day Starved","2 Days Starved"), 
                      values=c("black", "dimgrey", "lightgrey")) + 
  labs(x = "Head length (mm)", y = "Density") + 
  theme(legend.position = "bottom")

ggplot(pardat, aes(x = head_width * 25.4, group=factor(diet), colour=factor(diet))) + 
  geom_density() +  
  guides(alpha = FALSE) + 
  scale_colour_manual (name = "",
                       labels = c("Fed","1 Day Starved","2 Days Starved"), 
                       values = c("black", "dimgrey", "lightgrey")) + 
  labs(x = "Head width (mm)", y = "Density") + 
  theme(legend.position = "bottom")

ggplot(pardat, aes(x=body_length * 25.4, group=factor(diet), colour=factor(diet))) + 
  geom_density() +  
  guides(alpha = FALSE) + 
  scale_colour_manual (name = "",labels = c("Fed","1 Day Starved","2 Days Starved"), 
                       values = c("black", "dimgrey", "lightgrey")) + 
  labs(x = "Body length (mm)", y = "Density") + 
  theme(legend.position = "bottom")

ggplot(pardat, aes(x=body_width* 25.4, group=factor(diet), colour=factor(diet))) + 
  geom_density() +  
  guides(alpha = FALSE) + 
  scale_colour_manual (name = "",
                       labels = c("Fed","1 Day Starved","2 Days Starved"), 
                       values = c("black", "dimgrey", "lightgrey")) + 
  labs(x = "Body width (mm)", y = "Density") + 
  theme(legend.position = "bottom")
```
## MANOVA

Testing for effect of treatment (# of days starved) on body measurements

```{r}
dependent.vars<-(as.matrix(pardat[,9:10]))
summary(manova(dependent.vars ~ pardat$diet))
```
```{r}
pardat %>%
  filter(body_length != "" & body_width != "") %>%
  group_by(diet) %>%
  mutate(body_length_mm = body_length*25.4, body_width_mm = body_width*25.4) %>%
  summarize(meanbl = mean(body_length_mm), nbl = n(),  sdbl = sd(body_length_mm), sebl = sdbl / sqrt(nbl), meanbw = mean(body_width_mm), nbw = n(),  sdbw = sd(body_width_mm), sebw = sdbw / sqrt(nbw))
           
```

## Prepare data for TI analysis

```{r}
pardat$feign10 <- ifelse(pardat$feign == "Y", 1, 0)
pardat$time10 <- ifelse(pardat$time == "120", 1, 0)
pardat$instar <- ifelse(pardat$head_width >= 0.031 & pardat$head_length >= 0.018, 3, 2)
pardat_w_measures <- subset(pardat, pardat$instar == 3 | pardat$instar == 2)
pardat_w_measures$instar <- as.factor(pardat_w_measures$instar)
```

## Analysis of Y/N engaged in TI trait

```{r}
fit0 <- glm(feign10 ~ 1, data = pardat_w_measures, family = binomial)
fit1 <- glm(feign10 ~ diet, data = pardat_w_measures, family = binomial)
fit2 <- glm(feign10 ~ diet + instar, data = pardat_w_measures, family = binomial)
fit3 <- glm(feign10 ~ diet + diet:instar, data = pardat_w_measures, family = binomial)
```

```{r}
anova(fit0,fit1, test='Chisq')
anova(fit1,fit2, test='Chisq')
anova(fit1,fit3, test='Chisq')
```

```{r}
emm <- emmeans(fit1, "diet")
contrast(emm, adjust="bonferroni", method ="pairwise")
```


```{r}
pardat_w_measures %>%
ggplot(aes(x = diet, y = feign10)) + 
    stat_summary(aes(group = instar, color = instar), fun.y = "mean", geom = "line", 
                 position = position_dodge(.2), size = 1) +
  labs(y = "Percent TI",x = "Feeding group", color = "Instar") +
  scale_color_manual(name = "Instar", labels=c("2nd","3rd"), 
                    values = c("lightgrey", "dimgrey")) + 
  scale_x_discrete(breaks=c("1", "2", "3"),
                   labels=c("Fed", "1 Day Starved", "2 Days Starved"))

```


## Analysis of number of prods to initiate TI

```{r}
fit0 <- glm(prod ~ 1, data = pardat_w_measures, family = quasipoisson)
fit1 <- glm(prod ~ diet, data = pardat_w_measures, family = quasipoisson)
fit2 <- glm(prod ~ diet + instar, data = pardat_w_measures, family = quasipoisson)
fit3 <- glm(prod ~ diet + diet:instar, data = pardat_w_measures, family = quasipoisson)
```

```{r}
anova(fit0,fit1, test='Chisq')
anova(fit1,fit2, test='Chisq')
anova(fit1,fit3, test='Chisq')
```


```{r}
emm <- emmeans(fit1, "diet")
contrast(emm, adjust="bonferroni", method ="pairwise")
```
```{r}
pardat_w_measures %>%
  filter(feign10 == 1 ) %>%
ggplot(aes(x = diet, y = prod)) + 
  geom_boxplot(fill = "lightgrey") +
  labs(y = "Stimuli before TI",x = "Feeding group") +
  scale_x_discrete(breaks=c("1", "2", "3"),
                   labels=c("Fed", "1 Day Starved", "2 Days Starved"))

```

## Analysis of TI length 


```{r}
fit0 <- glm(time10 ~ 1, data = pardat_w_measures, family = binomial)
fit1 <- glm(time10 ~ diet, data = pardat_w_measures, family = binomial)
fit2 <- glm(time10 ~ instar, data = pardat_w_measures, family = binomial)
fit3 <- glm(time10 ~ diet:instar, data = pardat_w_measures, family = binomial)

anova(fit0,fit1, test='Chisq')
anova(fit0,fit2, test='Chisq')
anova(fit0,fit3, test='Chisq')
```

```{r}
pardat_w_measures %>%
  filter(feign10 == 1 ) %>%
ggplot(aes(x = diet, y = time10)) + 
    stat_summary(aes(group = diet), fun.y = "mean", geom = "bar", 
                 position = position_dodge(.2), size = 1) +
  labs(y = "Percent Long TI",x = "Feeding group", color = "Instar") +
  scale_color_manual(name = "Instar", labels=c("2nd","3rd"), 
                    values = c("lightgrey", "dimgrey")) + 
  scale_x_discrete(breaks=c("1", "2", "3"),
                   labels=c("Fed", "1 Day Starved", "2 Days Starved"))

```

## Calculate stats on molting


```{r}
n2 <- nrow(subset(pardat_w_measures, pardat_w_measures$instar == 2))
n3 <- nrow(subset(pardat_w_measures, pardat_w_measures$instar == 3))

paste("The number of second instar larvae was",n2)
paste("The number of third instar larvae was",n3)
```


```{r}
fed_3rds <- length(which(pardat_w_measures$instar == 3 & pardat_w_measures$diet == "1"))/length(which(pardat_w_measures$diet == "1"))

paste("Proportion of fed larvae molting is",fed_3rds)

starved1_3rds <- length(which(pardat_w_measures$instar == 3 & pardat_w_measures$diet == "2"))/length(which(pardat_w_measures$diet == "2"))
paste("Proportion of 1 day starved larvae molting is", starved1_3rds)

starved2_3rds <- length(which(pardat_w_measures$instar == 3 & pardat_w_measures$diet == "3"))/length(which(pardat_w_measures$diet == "3"))
paste("Proportion of 2 days starved larvae molting is", starved2_3rds)
```



## Heritability results

## Prepare data

```{r}
ped <-subset(dat, dat$generation =="P" | dat$generation =="O")
ped_columns <-c("ID","family_mother", "family_father")
ped <- ped[ped_columns]
for(x in 1:3) ped[,x]<-as.factor(ped[,x])


herit_dat<-subset(dat, dat$generation == "P" | dat$generation == "O")
herit_columns<-c("ID","feign")
herit_dat <- herit_dat[herit_columns]
names(herit_dat)[1]<-"animal"

```

## Animal model in MCMCglmm

```{r}
prior <- list(R = list(V = 1, fix = 1), 
              G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))


model <- MCMCglmm(feign ~ 1, random = ~animal, 
                  family = "categorical", 
                  prior = prior, pedigree = ped, data = herit_dat, 
                  nitt = 1000000, burnin = 50000, thin = 200, verbose=FALSE)
```

```{r}
herit <- model$VCV[, "animal"]/(model$VCV[, "animal"] +
      model$VCV[, "units"] + 1)

paste("The heritability estimate is", mean(herit))
HPDinterval(herit)
```

## Visualize heritability results

```{r}
herit_plot <- dat %>% group_by(family_ID, generation) %>% filter(generation!="N") %>% summarize(N = n(), feign_count = sum(feign == "Y"))
herit_plot$mean<-(herit_plot$feign_count/herit_plot$N)


offs<-subset(herit_plot,generation == "O")
off_cols <-c("family_ID","mean","N")
offs <- offs[off_cols]

herit_plot<-subset(herit_plot,generation == "P")

herit_plot<- merge(herit_plot,offs ,by = "family_ID")
herit_cols <-c("mean.x","mean.y","N.y")
herit_plot<-herit_plot[herit_cols]


jpeg("~/Desktop/Fig4.jpg", width = 20, height = 15, units = "cm", res = 300)

ggplot(herit_plot, aes(y = mean.y, x = as.factor(mean.x))) +
  geom_point(aes(size = N.y), alpha = 0.4) +
  labs(y = "Offspring mean", x = "Parent mean") + 
  theme(legend.position="none")
dev.off()
```

# Calculate sample sizes for figure legends

```{r}
pardat_w_measures %>%
  group_by(instar, diet) %>%
  summarize(n=n ())
```

```{r}
pardat_w_measures %>%
  filter(feign10 == 1) %>%
  group_by(diet) %>%
  summarize(n=n ())
```
